---
title: "PA1_template"
author: "JoAnna Tingey"
date: "Wednesday, May 13, 2015"
output: html_document
---
## Loading and preprocessing the data
```{r loadingdata}
    fileUrl="http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
    if(!file.exists("./repdata_data_activity.zip")) {download.file(fileUrl,destfile="./repdata_data_activity.zip",mode="wb")}
    if(!file.exists("./PA1Temp/activity.csv"))  {unzip(zipfile="./repdata_data_activity.zip",exdir="./PA1Temp")}
    repdata<-read.csv("./PA1Temp/activity.csv",header=TRUE,sep=",")
```


```{r ProcessData}
repdata$date<-as.Date(repdata$date,"%Y-%m-%d")
stepsbyday<-aggregate(steps~date, data=repdata,FUN="sum")
meanstepsbyinterval<-aggregate(steps~interval,data=repdata,FUN="mean")
```

## What is mean total number of steps taken per day?

```{r, TotalSteps}
hist(stepsbyday$steps,main = paste("Histogram of" , "Steps By Day"), xlab="Total Steps")
MeanSteps<-mean(stepsbyday$steps)
MedianSteps<-median(stepsbyday$steps)

```
The Total number of Steps per day has a Mean of `r MeanSteps` and Median of `r MedianSteps`.

## What is the average daily activity pattern?
```{r averagedaily}
plot(x=meanstepsbyinterval$interval,y=meanstepsbyinterval$steps, type="l", main=paste("Average Steps over time"), xlab="5 minute Interval", ylab="Average Steps")
max<-max(meanstepsbyinterval$steps)
minInterval<-subset(meanstepsbyinterval, steps==max)$interval
maxInterval<- minInterval+5
```
The 5 minute interval, on average across all the days in the dataset, contains the max number of steps is `r minInterval` - `r maxInterval`.

## Imputing missing values
```{r imputmissing}
row.has.na <- apply(repdata, 1, function(x){any(is.na(x))})
TotalNA<-nrow(repdata[row.has.na,])
for (i in which(is.na(repdata$steps))){
    repdata2<-repdata
    dat<-repdata[i,]
    mean<-subset(meanstepsbyinterval, interval==dat$interval )$steps
    repdata2[i,"steps"]<-mean
}
stepsbydayAdj<-aggregate(steps~date, data=repdata2,FUN="sum")
hist(stepsbydayAdj$steps,main = paste("Histogram of" , "Steps By Day"), xlab="Total Steps")
MeanStepsAdj<-mean(stepsbydayAdj$steps)
MedianStepsAdj<-median(stepsbydayAdj$steps)
```
Total NA's : `r TotalNA`
The Total number of Steps per day has a Mean of `r MeanStepsAdj` and Median of `r MedianStepsAdj`.
These are overall lower than the original Mean:`r MeanSteps` and Median:`r MedianSteps`

## Are there differences in activity patterns between weekdays and weekends?

```{r weekdayvsweekend}
repdata$factor<-ifelse(weekdays(repdata$date)==c("Sunday","Saturday"),"Weekend","Weekday")
repdata$factor<-as.factor(as.character(repdata$factor))
substepsbyweekday<-subset(repdata,factor=="Weekday")
substepsbyweekend<-subset(repdata,factor=="Weekend")
meanstepsbyintervalday<-aggregate(steps~interval,data=substepsbyweekday,FUN="mean")
meanstepsbyintervalend<-aggregate(steps~interval,data=substepsbyweekend,FUN="mean")
par(mfrow=c(1,1),mar=c(4,4,2,1))
plot(x<-meanstepsbyintervalday$interval,meanstepsbyintervalday[,"steps"],type='l',main="Weekday vs Weekend",ylab="Average Steps",xlab="")
points(x<-meanstepsbyintervalend$interval,meanstepsbyintervalend[,"steps"],type='l',col="red")
legend("topright", bty="n",legend = c("Weekday","Weekend"),col=c("black","red"),lwd=1)
```
